Analysis of NLA17 Dissolved Gas Indicator

### 2021-05-21 scripts/dgIndicatorAnalysis.Rmd

```{r}
# Read data file 
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

@@

@@

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()
```

@@

## Geopackage
Create a simple feature on entire dg data frame.  The geopackage will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
```{r, geopackage, eval=FALSE, echo=FALSE,}
library(janitor) # clean_names function
names(dg)

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm)

names(dg_1)
dg_1 <- clean_names(dg_1) %>% filter(ag.eco9.nm != "AK") %>% 
  mutate(log_n2o = round(log(dissolved_n2o),3)) %>% 
  mutate(log_n2osatratio = round(log(n2o_sat_ratio),3))
# underscores are "safer" names used in GIS

names(dg_1)
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # 4326 is just the EPSG identifier of WGS84. WGS84 comprises a standard coordinate frame for the Earth, a datum/reference ellipsoid for raw altitude data, and a gravitational equipotential surface (the geoid) that defines the nominal sea level.  5070 is for Albers equal Area.
st_crs(dg_sf1)
dim(dg_sf1) # 1184 observations

dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1184 obs vs. previously 1185 obs., why change in sample size
st_crs(dg_sf1)

@@

# write out dg.sf and ecoR as two layers in geopackage, gpkg, geodatabase
# 74. Creating a geopackage geodatabase, then adding more layers
# https://github.com/r-spatial/sf/issues/392
# creating a _v1.gpkg having large sample size and should be in Albers
st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
         delete_layer = TRUE) # needed, else function gives error if object already exists
st_layers("nladg_v1.gpkg")

# N.B. b/c old layer with untransformed was already in geodatabase have to delete it and then append the new layer
st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
st_layers("nla_dg1.gpkg")

@@

## Random Semivariograms
Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.  Downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.  An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.  Then pH values are randomly sampled and assigned to the coordinates so that a random semivariogram can be calculated.  That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram.

The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
# also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag.eco9.nm, ag.eco9.nm, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag.eco9.nm != "AK")

names(dg_1)

#filter in View returns 514 nitrate_n_result equal to zero
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
  mutate(log_ntl = round(log(ntl_lab_result),3)) %>% 
  mutate(log_ptl = round(log(ptl_lab_result),3)) %>% 
  mutate(log_chla = round(log(chla_result_vol),1)) %>% 
  mutate(log_si = round(log(silica_result),2)) %>% 
  mutate(log_ca = round(log(calcium_result),2)) %>% 
  mutate(ph_result = round(ph_result,2)) %>% 
  mutate(nitrate_n_b = case_when(
    nitrate_n_result <= 0.0001 ~ "low",
    nitrate_n_result > 0.0001 ~ "high"))

dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
summary(dg_1$nitrate_n_b)

@@

print(xyplot(gamma ~ dist, z_ph_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
))

z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
plot(variogram(ph_result~1, dg_sf1, cressie = TRUE))

print(xyplot(gamma ~ dist, z_cressie_ph, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1, cressie = TRUE)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
))

z_do_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
plot(z_do_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
plot(z_do_sf1, ylab = "NLA Surface DO Semivariance")

>>>>

```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

@@

names(dg)

@@

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()

@@

## Geopackage
Create a simple feature on entire dg data frame.  The geopackage will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
```{r, geopackage, eval=FALSE, echo=FALSE,}

names(dg)
dg_1 <- clean_names(dg)
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
dg_1 <- filter( dg_1, sample_source == "DG")

dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")

names(dg_1)
str(dg_1)

dg_1 %>% 
  ggplot(aes(as.factor(x = ag_eco9), y = log(dissolved_n2o_nmol))) + geom_boxplot()
# now no longer show a boxplot for NA category

dg_1 %>% 
  ggplot(aes(as.factor(x = ag_eco9), y = (n2o_sat_ratio))) + geom_boxplot()
# now no longer show a boxplot for NA category

#filter in View returns 514 nitrate_n_result equal to zero
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
  mutate(log_ntl = round(log(ntl_result),3)) %>% 
  mutate(log_ptl = round(log(ptl_result),3)) %>% 
  mutate(log_chla = round(log(chla_result),1)) %>% 
  mutate(log_si = round(log(silica_result),2)) %>% 
  mutate(log_ca = round(log(calcium_result),2)) %>% 
  mutate(ph_result = round(ph_result,2))
# %>% 
#   mutate(nitrate_n_b = case_when(
#     nitrate_n_result <= 0.0001 ~ "low",
#     nitrate_n_result > 0.0001 ~ "high"))  
# 
# dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
# summary(dg_1$nitrate_n_b)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting 
dim(dg_sf1) # 1184 observations

dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1082 obs vs. 1184 obs vs. previously 1185 obs., why change in sample size
st_crs(dg_sf1)

@@

# write out dg.sf and ecoR as two layers in geopackage, gpkg, geodatabase
# 74. Creating a geopackage geodatabase, then adding more layers
# https://github.com/r-spatial/sf/issues/392
# creating a _v1.gpkg having large sample size and should be in Albers
# st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
#          delete_layer = TRUE) # needed, else function gives error if object already exists
st_layers("nladg_v1.gpkg")

# N.B. b/c old layer with untransformed was already in geodatabase have to delete it and then append the new layer
st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
st_layers("nladg_v1.gpkg")

@@

## Random Semivariograms
Empirical semivariograms were calculated for:  pH, total nitrogen (ntl), total phosphorous (ptl), and chloropyll a (chla).  In addition, semivariograms were calculated for the dissolved gasses:  DO, N2O, CH4, and CO2. We downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.  An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.  Then pH values are randomly sampled and assigned to the coordinates so that a random semivariogram can be calculated.  That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram. Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.

The water quality variables of pH, tnl, tpl, and chla show strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.  The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation. The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.  CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)

# also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData

dg_1 <- filter( dg_1, sample_source == "DG")
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")

names(dg_1)
str(dg_1)

@@

print(xyplot(gamma ~ dist, z_ph_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
))

# used Cressie's robust variogram b/c skewness and not wanting to transform pH as it is already on a log scale
z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
plot(variogram(ph_result~1, dg_sf1, cressie = TRUE))

print(xyplot(gamma ~ dist, z_cressie_ph, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1, cressie = TRUE)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
))

## NTL
summary(dg_sf1$ntl_result) #no NA's

filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(ntl_result)) + geom_histogram()

# log transformation ~ unimodal & symmetric
filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(log(ntl_result))) + geom_histogram()

# hist(dg_sf1$ntl_result)
# hist(log(dg_sf1$ntl_result))

# default cutoff and width and no transformation
z_lntl_sf1 = variogram(log(ntl_result) ~ 1, dg_sf1)
plot(z_lntl_sf1, ylab = "NLA NTL Semivariance")
# set ylim for random semivariograms based on this plot

# cutoff and width from King et al. paper still not working even with tranform
z1_ntl_king = variogram(log(ntl_result) ~1, dg_sf1, cutoff=2500, width=20)
plot(z1_ntl_king)

# plot(z_ph_sf1, plot.numbers = TRUE)
# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")

print(xyplot(gamma ~ dist, z_lntl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ntl_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NTL semivariance'
))

## PTL
summary(dg_sf1$ptl_result) # no NA's

filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(ptl_result)) + geom_histogram()

# log transformation ~ unimodal & symmetric
filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(log(ptl_result))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lptl_sf1 = filter(dg_sf1, (!is.na(ptl_result))) %>% variogram(log(ptl_result)~1, .)
plot(z_lptl_sf1, ylab = "NLA TP Semivariance")
# set ylim of random semivariograms based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(ptl_result)))

print(xyplot(gamma ~ dist, z_lptl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ptl_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2), xlab = 'distance', ylab = 'TP semivariance'
))

## CHLA
summary(dg_sf1$chla_result) # no NA's

filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(chla_result)) + geom_histogram()

# log transformation not that great but use for now
filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(log(chla_result))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lchla_sf1 = filter(dg_sf1, (!is.na(chla_result))) %>% variogram(log(chla_result)~1, .)
plot(z_lchla_sf1, ylab = "NLA CHLA Semivariance")
# set ylim of random semivariograms based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(chla_result)))

print(xyplot(gamma ~ dist, z_lchla_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$chla_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2.5), xlab = 'distance', ylab = 'CHLa semivariance'
))

## DO
summary(dg_sf1$o2_surf) # have 4 NA's

filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(o2_surf)) + geom_histogram()

# log probably too strong of transformation but use for now
filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(log(o2_surf))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_ldo_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
# plot(z_ldo_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
plot(z_ldo_sf1, ylab = "NLA Surface DO Semivariance")

# data frame with no NA's
dg_sf2 <- filter(dg_sf1, (!is.na(o2_surf)))

print(xyplot(gamma ~ dist, z_ldo_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf2$random = sample(dg_sf2$o2_surf)
                     v = variogram(log(random) ~ 1, dg_sf2)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.12), xlab = 'distance', ylab = 'NLA DO semivariance'
))

## N2O
summary(dg_sf1$dissolved_n2o_nmol) # no NA's

filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()

# unimodal, but not symmetrical
filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()

z_n2o_sf1 <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol) ~1, .)
plot(z_n2o_sf1, ylab = "N2O Semivariance")

# change cutoff and width
z_n2o = filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol)~1, ., cutoff = 750000, width = 25000)
# plot(z_n2o, ylab = "N2O Semivariance", plot.numbers = TRUE)
plot(z_n2o, ylab = "NLA N2O Semivariance")

print(xyplot(gamma ~ dist, z_n2o, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
    panel = function(x, y, ...) {
        for (i in 1:100) {
            dg_sf1$random = sample(dg_sf1$dissolved_n2o_nmol)
            v = variogram(log(random) ~ 1, dg_sf1, cutoff = 750000, width = 25000)
            llines(v$dist, v$gamma, col = 'grey')
        }
        panel.xyplot(x, y, ...)
    },
    ylim = c(0, 0.35), xlab = 'distance', ylab = 'N2O semivariance'
))

## CH4
summary(dg_1$dissolved_ch4_umo) # no NA's

filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(dissolved_ch4_umol)) + geom_histogram()

# log transformation looks good, ~ unimodal & symmetric
filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(log(dissolved_ch4_umol))) + geom_histogram()

z_lch4_sf1 = filter(dg_sf1, (!is.na(dissolved_ch4_umol))) %>% variogram(log(dissolved_ch4_umol)~1, .)
plot(z_lch4_sf1, ylab = "NLA CH4 Semivariance")
# change ylim below in random semivariogram based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's for random sampling below
# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_ch4_umol)))

print(xyplot(gamma ~ dist, z_lch4_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$dissolved_ch4_umol)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2), xlab = 'distance', ylab = 'NLA CH4 semivariance'
))

## CO2
summary(dg_1$dissolved_co2_umo) # no NA's

filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(dissolved_co2_umol)) + geom_histogram()

# log transformation, unimodal and symmetric
filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(log(dissolved_co2_umol))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lco2_sf1 = filter(dg_sf1, (!is.na(dissolved_co2_umol))) %>% variogram(log(dissolved_co2_umol)~1, .)
plot(z_lco2_sf1, ylab = "NLA CO2 Semivariance")
# change ylim in random semivariogram based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_co2_umol)))

print(xyplot(gamma ~ dist, z_lco2_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$dissolved_co2_umol)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NLA CO2 semivariance'
))

## Random Ecoregional Semivariogram N2O for ASLO Abstract
I used ecoregional-specific data frames to feed into random semivariogram function.  
```{r, ecoregion_semivar}

## N2O
summary(dg_sf1$dissolved_n2o_nmol) # no NA's

filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()

# unimodal, but not symmetrical
filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER") %>%  ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()

# zcpl <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="CPL")
# zcpl_variog <- variogram(log(dissolved_n2o_nmol) ~1, zcpl)
# plot(zcpl_variog, ylab = "N2O zCPL Semivariance")
# the 2-step process above matches the piped process below
xer_n2o_sf1 <- filter(dg_sf1,!is.na(dissolved_n2o_nmol)) %>%
  filter(ag_eco9 =="XER") %>% 
  variogram(log(dissolved_n2o_nmol) ~1, .)
plot(xer_n2o_sf1, ylab = "N2O XER Semivariance")

xer <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER")

print(xyplot(gamma ~ dist, wmt_n2o_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
    panel = function(x, y, ...) {
        for (i in 1:100) {
            xer$random = sample(xer$dissolved_n2o_nmol)
            v = variogram(log(random) ~ 1, xer)
            llines(v$dist, v$gamma, col = 'grey')
        }
        panel.xyplot(x, y, ...)
    },
    ylim = c(0, 0.35), xlab = 'distance', ylab = 'XER N2O semivariance'
))
